---
title: "08.compare_mce_maier"
output: github_document
date: "2024-10-02"
---  
  
In this file we prepare the final figures from the data analysed in `08.compare_mce_maier.ipynb`. 

```{r setup, include=FALSE}
library(tidyverse)
library(ggrepel)
library(readxl)
```

## Prepare directories. 
  
```{r prep.dirs}

OUTPUT_DIR <- "../data/08.compare_mce_maier/"

```


## Read data.  
  
Here we read the UMAP coordinates.  
```{r read.data}

# UMAP based on MolE representation
mole.umap <- read_tsv(file.path(OUTPUT_DIR, "mole_joint_umap.tsv.gz"), 
                      show_col_types = FALSE)

# UMAP based on ECFP4 representation
ecfp4.umap <- read_tsv(file.path(OUTPUT_DIR, "ecfp4_joint_umap.tsv.gz"), 
                      show_col_types = FALSE)


# Complete prediction set
completepreds <- read_excel("../data/04.new_predictions/mole_mce_predictions_litsearch.xlsx", sheet = "mole_prediction_overview")

# Literature search data. 
litsearch.mole <- read_excel("../data/04.new_predictions/mole_mce_predictions_litsearch.xlsx", sheet = "mole_over10")

mce_litexamples <- litsearch.mole %>% 
  filter(!is.na(`Reported Activity`),
         antibiotic == "not_abx") %>% 
  select(`Catalog Number`, `Reported Activity`) %>% 
  rename("chem_id" = "Catalog Number")

mce_litexamples %>% head()
```
  
## Plot UMAP. 

Plot the UMAP, highlighting some chemicals that are predicted to be broad-spectrum and confirmed in the literature.

```{r chems.interest}
# Chemicals of interest

coi_df <- data.frame("chem_id" = c("HY-B0183", "HY-B0021", "HY-N0829", "HY-B0723", "HY-B2136" ),
                     "ProductName" = c("Ellagic acid", "Doxifluridine", "Shionone", "Ospemifene", "Tannic acid"))

coi_df
```

```{r}

mce_litexamples.umap <- mce_litexamples %>% 
  left_join(mole.umap, by="chem_id") %>% 
  mutate(`Chemical Library` = "MCE (Predictions with reported activity)")
```


```{r}
coi.mole <- coi_df %>% 
  left_join(mole.umap, by="chem_id")
```
 
```{r plot.umap.mole}
mole.umap <- mole.umap %>% 
  mutate(keep = case_when(`Chemical Library` == "Maier et.al." ~ TRUE,
                          `Chemical Library` == "MCE" & chem_id %in% completepreds$`Catalog Number` ~ TRUE,
                          TRUE ~ FALSE)) %>% 
  filter(keep)



u.mole <- ggplot(mole.umap, aes(x=umap1, y=umap2, fill=`Chemical Library`)) +
  
  
  geom_point(size=2, color="white", shape=21, alpha=0.75) +
  geom_point(data = mce_litexamples.umap, color="white", size=2.5, shape=21, alpha=0.75) +
  
  
  geom_text_repel(data = coi.mole, aes(label=ProductName), 
                  max.overlaps = Inf, size=3, min.segment.length = 0, 
                  color="black", fontface="plain",
                  box.padding = 0.5,
                  nudge_x = if_else(coi.mole$ProductName %in% c("Ellagic acid", "Tannic acid", "Ospemifene"), 1, 0),
                  
                  nudge_y = case_when(coi.mole$ProductName %in% c("Ospemifene", "Shionone") ~ 3,
                                      coi.mole$ProductName %in% c("Doxifluridine") ~ -2,
                                      coi.mole$ProductName %in% c("Tannic acid") ~ 2,
                                      TRUE ~ 0)) +
  
  
  scale_fill_manual(values=c("#DE1F84", "#C5C5C5", "#1F9DBB")) +
  
  theme_bw() +
  
  theme(legend.position = "bottom",
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_rect(fill = "transparent", color=NA),
      plot.background = element_rect(fill = "transparent", colour = NA))  + 
  
labs(x="UMAP 1",
       y="UMAP 2")

u.mole
ggsave(filename = file.path(OUTPUT_DIR, "mole_joint_umap.pdf"), plot=u.mole, dpi = 300,
       height = 15, width = 21, units="cm")

ggsave(filename = file.path(OUTPUT_DIR, "mole_joint_umap.png"), plot=u.mole, dpi = 300,
       height = 15, width = 21, units="cm")
  

```


```{r}
mce_metadata <- completepreds %>% 
  select(`Catalog Number`, ProductName,  nk_total, antibiotic) %>% 
  rename("chem_id" = "Catalog Number") %>% 
  
  left_join(mce_litexamples, by="chem_id")
  

mole.umap %>% 
  left_join(mce_metadata, by="chem_id") %>% 
  select(-keep) %>% 
  rename("n_predicted_inhibited_strains" = "nk_total") %>% 
  write_tsv("../data/08.compare_mce_maier/umap_library_comparison.tsv")
```








The same but for ECFP4. 
```{r plot.umap.fps}

mce_litexamples.umap <- mce_litexamples %>% 
  left_join(ecfp4.umap, by="chem_id") %>% 
  mutate(`Chemical Library` = "MCE (Predictions with reported activity)")

coi.ecfp4 <- coi_df %>% 
  left_join(ecfp4.umap, by="chem_id")



u.ecfp4 <- ggplot(ecfp4.umap, aes(x=umap1, y=umap2, fill=`Chemical Library`)) +
  geom_point(size=1.5, color="white", shape=21, alpha=0.75) +
  geom_point(data = mce_litexamples.umap, color="white", size=2.5, shape=21, alpha=0.75) +
  
  
  geom_text_repel(data = coi.ecfp4, aes(label=ProductName), 
                  max.overlaps = Inf, size=3, min.segment.length = 0, 
                  color="black", fontface="plain",
                  box.padding = 0.5,
                   nudge_y = if_else(coi.mole$ProductName %in% c("Ospemifene", "Tannic acid"), -5, 0),
                  nudge_x = if_else(coi.mole$ProductName %in% c("Doxifluridine", "Ellagic acid"), -3, 0)) +
  
  
  scale_fill_manual(values=c("#DE1F84", "#C5C5C5","#1F9DBB")) +
  
  theme_bw() +
  
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_rect(fill = "transparent", color=NA),
      plot.background = element_rect(fill = "transparent", colour = NA))  + 
  
labs(x="UMAP 1",
       y="UMAP 2")

u.ecfp4
ggsave(filename = file.path(OUTPUT_DIR, "ecfp4_joint_umap.pdf"), plot=u.ecfp4, dpi = 300,
       height = 13, width = 19, units="cm")
  

```

## Comparing MolE and ECFP4

```{r panel.lib}
library(ggpubr)
```


```{r panel}

umap.arranged <- ggarrange(u.mole + ggtitle("MolE"), 
          u.ecfp4 + ggtitle("ECFP4"), 
          common.legend = TRUE,
          legend="bottom")

umap.arranged
ggsave(filename = file.path(OUTPUT_DIR, "joint_umap_panel.pdf"), plot=umap.arranged, dpi = 300,
       height = 11, width = 20, units="cm")
```
  
## Comparing molecular similarity. 
  
```{r read.distance.qs}

# Read molecular similarity
mol.similarity <- read_tsv("../data/08.compare_mce_maier/ecfp4_distance_to_maier.tsv.gz", show_col_types = FALSE) %>% 
  # Make tidy form to ease handling
  pivot_longer(cols = -selected_chem, names_to = "prestwick_ID", values_to = "jaccard_distance") %>% 
  
  # Get tanimoto similarity
  mutate(tanimoto_similarity = 1-jaccard_distance) %>% 

  # Gather the ranking per molecule. Make sure to rank ties the same
  group_by(selected_chem) %>% 
  mutate(similarity_ranking = rank(-tanimoto_similarity, ties.method = "min"))

# Read prestwick metadata
maier_chems_metadata = read_excel("../raw_data/maier_microbiome/chem_library_info_SF1.xlsx") %>% 
  select(prestwick_ID, `chemical name`, `target species`) %>% 
  rename("prestwick_chemname" = "chemical name",
         "target_species" = "target species")


# Read screening results
maier_screening_results <- read_tsv("../data/01.prepare_training_data/maier_screening_results.tsv.gz", show_col_types = FALSE) %>% 
  # Gather chemicals used in training
  filter(prestwick_ID %in% unique(mol.similarity$prestwick_ID)) %>% 
  
  # Count number of strains inhibited
  pivot_longer(cols = -prestwick_ID, names_to = "strain", values_to = "growth_inhibition") %>% 
  group_by(prestwick_ID) %>% 
  summarise(n_inhibited_strains = sum(growth_inhibition)) %>% 
  ungroup()


# Join information
mol.similarity <- mol.similarity %>% 
  left_join(maier_chems_metadata, by="prestwick_ID") %>% 
  left_join(maier_screening_results, by="prestwick_ID")

```

  
Determine the most similar molecule percategory
```{r}
# Overall
overall.most_similar <- mol.similarity %>% 
  filter(similarity_ranking == 1) %>% 
  # Thymidine has two molecules with equal ranking
  head(6) %>% 
  
  mutate(sim_category = "Overall")

# Most similar antibiotic
abx.most_similar <- mol.similarity %>% 
  filter(target_species == "bacteria") %>% 
  group_by(selected_chem) %>% 
  filter(similarity_ranking == min(similarity_ranking)) %>% 
  mutate(sim_category = "Antibiotic")

# Most similar non-antibiotic with broad spectrum activity
nonabx.bsa.most_similar <- mol.similarity %>% 
  filter(target_species != "bacteria",
         n_inhibited_strains >= 10) %>% 
  group_by(selected_chem) %>% 
  filter(similarity_ranking == min(similarity_ranking)) %>% 
  mutate(sim_category = "Non-antibiotic with BSA")


most_similar.df <- bind_rows(overall.most_similar, abx.most_similar, nonabx.bsa.most_similar) %>% 
  select(selected_chem, prestwick_ID, tanimoto_similarity, similarity_ranking, target_species, n_inhibited_strains, sim_category) %>% 
  
  # For some compounds, the same molecule is the most similar in more than one category. Let's choose the more specific annotation.
  group_by(selected_chem, prestwick_ID) %>% 
  slice_tail() %>% 
  
  # Prepare labelling
  mutate(tansim_label = round(tanimoto_similarity, 2),
         sim_category = factor(sim_category, levels=c("Overall", "Antibiotic", "Non-antibiotic with BSA")))
  
  
most_similar.df
```


```{r}
similarity.chosen <- ggplot(mol.similarity, aes(x=similarity_ranking, y=tanimoto_similarity)) +
  geom_hline(yintercept = 0.3, linetype="longdash") +
  
  geom_point(size=1.8, color="white", shape=21, alpha=1, fill="#C5C5C5") +
  geom_point(data=most_similar.df, aes(fill=sim_category, shape=sim_category), size=2, 
             color="black", shape=21, alpha=0.75) +
  
  scale_fill_manual(values=c("#377eb8", "#e41a1c", "#1b9e77")) +
  
  
  geom_text_repel(data = most_similar.df, aes(label=tansim_label), 
                  max.overlaps = Inf, size=3.5, 
                  min.segment.length = 0, 
                  color="black", 
                  fontface="plain",
                  point.padding = 0.2,
                  box.padding = 0.2,
                  segment.size=0.3,
                  
                  nudge_y = case_when(most_similar.df$selected_chem == "Cetrorelix" & most_similar.df$sim_category == "Non-antibiotic with BSA" ~ -0.125,
                                      most_similar.df$selected_chem == "Cetrorelix" & most_similar.df$sim_category == "Overall" ~ 0.125,
                                      most_similar.df$selected_chem == "Cetrorelix" & most_similar.df$sim_category == "Antibiotic" ~ -0.05,
                                      
                                      most_similar.df$selected_chem == "Elvitegravir" & most_similar.df$sim_category == "Non-antibiotic with BSA" ~ -0.15,
                                      most_similar.df$selected_chem == "Elvitegravir" & most_similar.df$sim_category == "Antibiotic" ~ 0.1255,
                                      
                                      most_similar.df$selected_chem == "Opicapone" & most_similar.df$sim_category == "Antibiotic" ~ 0.03,
                                      most_similar.df$selected_chem == "Opicapone" & most_similar.df$sim_category == "Non-antibiotic with BSA" ~ 0.15,
                                      
                                      most_similar.df$selected_chem == "Visomitin" & most_similar.df$sim_category == "Overall" ~ 0.15,
                                      most_similar.df$selected_chem == "Visomitin" & most_similar.df$sim_category == "Antibiotic" ~ -0.125,
                                      
                                      most_similar.df$selected_chem == "Ebastine" & most_similar.df$sim_category == "Antibiotic" ~ -0.15,
                                      most_similar.df$selected_chem == "Ebastine" & most_similar.df$sim_category == "Overall" ~ 0.125,
                                      
                                      TRUE ~ 0),
                  
                  nudge_x = case_when(most_similar.df$selected_chem == "Ebastine" & most_similar.df$sim_category == "Non-antibiotic with BSA" ~ 125,
                                      most_similar.df$selected_chem == "Ebastine" & most_similar.df$sim_category == "Overall" ~ 125,
                                      
                                      most_similar.df$selected_chem == "Thymidine" & most_similar.df$sim_category == "Non-antibiotic with BSA" ~ 150,
                                      most_similar.df$selected_chem == "Thymidine" & most_similar.df$sim_category == "Antibiotic" ~ 150,
                                      
                                      most_similar.df$selected_chem == "Opicapone" & most_similar.df$sim_category == "Antibiotic" ~ 150,
                                      
                                      most_similar.df$selected_chem == "Elvitegravir" & most_similar.df$sim_category == "Antibiotic" ~ 100,
                                      
                                      most_similar.df$selected_chem == "Visomitin" & most_similar.df$sim_category == "Non-antibiotic with BSA" ~ 180,
                                      
                                      most_similar.df$selected_chem == "Cetrorelix" & most_similar.df$sim_category == "Antibiotic" ~ 125,
                                      TRUE ~ 0)) +
  
  
  facet_wrap(~selected_chem) +
  
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(size=6)) +
  labs(x="Ranking of training set compounds",
       y="Tanimoto similarity",
       fill = "Similarity category")


ggsave(plot=similarity.chosen, filename = "../data/08.compare_mce_maier/similarity_to_maier.png",
       dpi = 300, width = 21, height = 12.5, units="cm")

ggsave(plot=similarity.chosen, filename = "../data/08.compare_mce_maier/similarity_to_maier.pdf",
       dpi = 300, width = 21, height = 12.5, units="cm")
```



```{r}
most_similar_metadata <- most_similar.df %>% 
  unite("selected_prest", c(selected_chem, prestwick_ID), sep=":") %>% 
  select(selected_prest, sim_category)
```




```{r}

mol.similarity %>% 
  unite("selected_prest", c(selected_chem, prestwick_ID), sep=":", remove=FALSE) %>% 
  
  left_join(most_similar_metadata, by="selected_prest") %>% 
  select(-selected_prest) %>% 
  
  arrange(selected_chem, similarity_ranking) %>% 
  
  rename("measured_n_inhibited_strains" = "n_inhibited_strains",
         "most_similar_category" = "sim_category") %>%
  write_tsv("../data/08.compare_mce_maier/selected_prestwick_tanimoto.tsv")
  
  

```

